---
title: "Get started with ccldr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with ccldr}
  %\VignetteEngine{knitr::rmarkdown_notangle}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE,
  purl = FALSE
)
```

`ccldr` is built for analysts who need California Community Care Licensing
Division (CCLD) facility data in R without leaving their scripts. It focuses on
three workflows:

- verify license numbers;
- fetch full detail for one facility;
- build current Alameda County child-care snapshots;
- append Census geographies to addressable facility rows.

The examples are not evaluated when the vignette is built because they call the
live CCLD Transparency API. Representative output is shown below the main
snippets so you can see the result shape before running them.

## Verify a column

Use `ccld_verify()` when you have a vector or a data-frame column of possible
license numbers. The CCLD API expects facility numbers in a 9-digit form, but
`ccld_verify()` accepts 8- or 9-digit inputs and pads internally before making
requests. It returns one row per input, including unknown licenses.

```{r}
library(ccldr)
library(dplyr)
library(readr)

rr <- read_csv("rr_sites.csv")

verified <- ccld_verify(rr$license_number)

rr |>
  left_join(verified, by = c("license_number" = "input")) |>
  select(site_name, license_number, found, facility_name, status)
```

Example output:

```text
#> # A tibble: 2 x 5
#>   site_name                 license_number found facility_name       status
#>   <chr>                     <chr>          <lgl> <chr>               <chr>
#> 1 Johnson Family Child Care 13423996       TRUE  JOHNSON III, JOHNNY Licensed
#> 2 Unknown Site              99999999       FALSE <NA>                <NA>
```

Keeping `found = FALSE` rows in the result makes audits easier: your original
records stay visible instead of disappearing during a join.

## Inspect one facility

Use `ccld_facility()` when you need the full API record for one known license.
Like `ccld_verify()`, it accepts either the 8- or 9-digit license form. The
result is a one-row tibble with scalar facility fields and nested list-columns
for reports and complaints.

```{r}
facility <- ccld_facility("13423996")

facility |>
  select(facility_name, status, capacity, last_visit_date)
```

Example output:

```text
#> # A tibble: 1 x 4
#>   facility_name       status   capacity last_visit_date
#>   <chr>               <chr>       <int> <date>
#> 1 JOHNSON III, JOHNNY Licensed       14 2024-11-04
```

```{r}
facility |>
  tidyr::unnest(reports) |>
  select(facility_number, report_date, report_type)
```

Example output:

```text
#> # A tibble: 1 x 3
#>   facility_number report_date report_type
#>   <chr>           <date>      <chr>
#> 1 013423996       2024-11-04  Other
```

```{r}
facility |>
  tidyr::unnest(complaints)
```

Example output:

```text
#> # A tibble: 0 x 4
#> # i 4 variables: facility_number <chr>, complaint_date <date>,
#> #   allegation <chr>, outcome <chr>
```

`ccld_facility()` is scalar by design. For a column of candidate licenses, run
`ccld_verify()` first. Its slim output includes `date_closed` for closure
audits, and you can request full detail only for the facilities where you need
visits, complaint counts, reports, or itemized complaints.

## Pull an Alameda snapshot

Use `ccld_alameda()` when you need the current set of Alameda County facilities
for a supported child-care type.

```{r}
preschools <- ccld_alameda("preschools")
large_fccs <- ccld_alameda("large_fccs")

preschools |>
  select(facility_number, facility_name, status, city, zip) |>
  slice_head(n = 3)
```

Example output:

```text
#> # A tibble: 3 x 5
#>   facility_number facility_name          status   city    zip
#>   <chr>           <chr>                  <chr>    <chr>   <chr>
#> 1 013423751       WILD CHILD SCHOOLHOUSE Licensed Oakland 94611
#> 2 013423056       AGAPE LEARNING CENTER  Licensed Oakland 94621
#> 3 013422467       COLIBRI PRESCHOOL      Licensed Oakland 94611
```

Supported values are:

- `"large_fccs"`
- `"infant_centers"`
- `"school_age_centers"`
- `"preschools"`
- `"single_licensed_centers"`

`"small_fccs"` is not supported because the API blocks that search route.
`"centers"` is not supported because the API bucket is sparse and legacy; use
`"preschools"` for the center workflow.

## Add Census geographies

Use `ccld_add_census_geographies()` when you need coordinates or Census
geography identifiers for facility rows with addresses. The helper uses the
U.S. Census Geocoder, does not require an API key, deduplicates repeated
addresses, and caches responses on disk.

```{r}
preschools_geo <- preschools |>
  ccld_add_census_geographies()

preschools_geo |>
  select(
    facility_number,
    geocode_status,
    latitude,
    longitude,
    census_tract_geoid,
    census_block_geoid,
    zcta_geoid,
    unified_school_district_name
  ) |>
  slice_head(n = 3)
```

Example output:

```text
#> # A tibble: 3 x 8
#>   facility_number geocode_status latitude longitude census_tract_geoid
#>   <chr>           <chr>             <dbl>     <dbl> <chr>
#> 1 013423751       matched            37.8     -122. 06001401100
#> 2 013423056       matched            37.7     -122. 06001409400
#> 3 013422467       matched            37.8     -122. 06001401100
#> # i 3 more variables: census_block_geoid <chr>, zcta_geoid <chr>,
#> #   unified_school_district_name <chr>
```

## Cache behavior

Every API response is cached on disk for 24 hours by default.

```{r}
ccld_cache_info()
ccld_cache_clear()
```

`ccld_cache_info()` returns a tibble with cache file paths, ages, and sizes.
`ccld_cache_clear()` returns the number of cached files removed.

For a fresh response in one call, set `cache = FALSE`:

```{r}
ccld_verify("13423996", cache = FALSE)
```

For debugging, these options are useful:

```{r}
options(ccldr.verbose = TRUE)
options(ccldr.delay = 1)
options(ccldr.cache_ttl_seconds = 3600)
```

Read the website articles for recurring script patterns and troubleshooting:

- <https://chekos.github.io/ccldr/articles/closure-date-audits.html>
- <https://chekos.github.io/ccldr/articles/live-api-workflows.html>
- <https://chekos.github.io/ccldr/articles/troubleshooting.html>
