---
title: "Getting started with ccldr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with ccldr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

`ccldr` is a tidyverse-native R client for the CCLD Transparency API. This
vignette walks through the three things it is built to do.

## Verify a column of license numbers

Imagine you have a CSV of new R&R intakes with a `license_number` column, and
you want to know which licenses CCLD recognises.

```{r}
library(ccldr)
library(dplyr)
library(readr)

rr <- read_csv("rr_sites.csv")
verified <- ccld_verify(rr$license_number)

rr |>
  left_join(verified, by = c("license_number" = "input")) |>
  filter(!found) |>
  select(license_number, site_name)
```

`ccld_verify()` returns one row per input. Unknown licenses do not drop; they
appear with `found = FALSE` and `NA`s in the data columns.

## Audit a specific site

When you need every field CCLD has for one license, use `ccld_facility()`.

```{r}
f <- ccld_facility("13423996")

f |> select(visits_total, visits_complaints, last_visit_date)
f |> tidyr::unnest(complaints)
f |> tidyr::unnest(reports)
```

## Live Alameda snapshot

When you need the current population of a facility type in Alameda, use
`ccld_alameda()`.

```{r}
preschools <- ccld_alameda("preschools")
large_fccs <- ccld_alameda("large_fccs")
```

Each call walks the 17 Alameda cities under the hood because the API's
`FacilitySearch` endpoint caps each response at 250 results.

There is no `"small_fccs"` option because the API blocks `facType = 0`
searches. For small FCC enumeration, use the open-data CSVs at
[`chekos/ccld-open-data-snapshot`](https://github.com/chekos/ccld-open-data-snapshot).

## Caching

Every API response is cached on disk for 24 hours by default. To bypass the
cache for one call, pass `cache = FALSE`. To clear the whole cache, call
`ccld_cache_clear()`. To inspect it, call `ccld_cache_info()`.
