---
title: "Capacity by geography"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE,
  purl = FALSE
)
```

Use this workflow when you want a small analytical report from live CCLD data:
where licensed capacity is concentrated, how that varies by city or Census
geography, and which closed records should be separated from current supply.

The examples are not evaluated during site builds because they call the live
CCLD Transparency API and the Census Geocoder.

## Build an analysis file

Start with one Alameda snapshot, then enrich it with the fields that only come
from the facility detail endpoint. This example uses preschools, but the same
pattern works for the other supported `ccld_alameda()` facility types.

```{r}
library(ccldr)
library(dplyr)
library(ggplot2)

preschools <- ccld_alameda("preschools")

verified <- ccld_verify(preschools$facility_number) |>
  select(
    facility_number,
    verified_found = found,
    verified_status = status,
    verified_city = city,
    capacity,
    date_closed,
    license_effective_date,
    last_visit_date
  )

preschools_detail <- preschools |>
  select(-capacity, -date_closed, -license_effective_date, -last_visit_date) |>
  left_join(verified, by = "facility_number") |>
  mutate(
    found = coalesce(verified_found, found),
    status = coalesce(verified_status, status),
    city = coalesce(city, verified_city),
    is_closed = !is.na(date_closed) |
      grepl("^Closed", status, ignore.case = TRUE),
    is_open = found & !is_closed & !is.na(capacity)
  ) |>
  select(-verified_found, -verified_status, -verified_city)
```

The Alameda search endpoint is useful for finding records, but it does not
return `capacity` or `date_closed`. Calling `ccld_verify()` for the facility
numbers fills those fields from `FacilityDetail`.

## Add Census geographies

Append coordinates and geography identifiers before summarizing. Keep unmatched
geocodes in the analysis table so missing addresses remain visible.

```{r}
preschools_geo <- preschools_detail |>
  ccld_add_census_geographies()

preschools_geo |>
  count(geocode_status, sort = TRUE)
```

For a city-only report, you can skip the geocoding step and use
`preschools_detail` directly. For ZCTA, tract, block group, school district, or
legislative district summaries, use the geocoded table.

## Capacity by city

Treat open facilities as current supply. Closed facilities are still valuable
for audits, but mixing them into current capacity totals will overstate supply.

```{r}
open_preschools <- preschools_geo |>
  filter(is_open)

city_capacity <- open_preschools |>
  group_by(city) |>
  summarise(
    facilities = n(),
    total_capacity = sum(capacity),
    median_capacity = median(capacity),
    capacity_per_facility = total_capacity / facilities,
    .groups = "drop"
  ) |>
  arrange(desc(total_capacity))

city_capacity
```

```{r}
ggplot(city_capacity, aes(x = reorder(city, total_capacity), y = total_capacity)) +
  geom_col(fill = "#2b8cbe") +
  coord_flip() +
  labs(
    title = "Total licensed preschool capacity by city",
    subtitle = "Open facilities only",
    x = NULL,
    y = "Licensed capacity"
  ) +
  theme_minimal()
```

Look at both total capacity and facility count. A city can have a high total
because it has many small sites, a few large sites, or both.

```{r}
ggplot(city_capacity, aes(x = facilities, y = total_capacity)) +
  geom_point(aes(size = median_capacity), alpha = 0.7, color = "#225ea8") +
  geom_text(aes(label = city), check_overlap = TRUE, nudge_y = 25, size = 3) +
  scale_size_continuous(name = "Median capacity") +
  labs(
    title = "Facility count and licensed capacity move together, but not perfectly",
    x = "Open facilities",
    y = "Total licensed capacity"
  ) +
  theme_minimal()
```

## Capacity by ZCTA

ZCTAs are useful for a first neighborhood-scale scan because they are familiar
to many partners and are returned by the Census Geocoder.

```{r}
zcta_capacity <- open_preschools |>
  filter(!is.na(zcta_geoid)) |>
  group_by(zcta_geoid) |>
  summarise(
    facilities = n(),
    total_capacity = sum(capacity),
    median_capacity = median(capacity),
    cities = paste(sort(unique(city)), collapse = ", "),
    .groups = "drop"
  ) |>
  arrange(desc(total_capacity))

zcta_capacity |>
  slice_head(n = 15)
```

```{r}
zcta_capacity |>
  slice_max(total_capacity, n = 15) |>
  ggplot(aes(x = reorder(zcta_geoid, total_capacity), y = total_capacity)) +
  geom_col(fill = "#756bb1") +
  coord_flip() +
  labs(
    title = "Highest-capacity ZCTAs",
    subtitle = "Open preschool records with matched Census geographies",
    x = "ZCTA",
    y = "Licensed capacity"
  ) +
  theme_minimal()
```

Capacity totals are not access rates. To turn this into an access analysis,
join child population, poverty, language, or other denominator data to the same
ZCTA or tract identifiers.

## Capacity by tract

Census tracts give a finer-grained view than ZCTAs. Use them to find local
clusters, then inspect the underlying facilities before drawing conclusions.

```{r}
tract_capacity <- open_preschools |>
  filter(!is.na(census_tract_geoid)) |>
  group_by(census_tract_geoid) |>
  summarise(
    facilities = n(),
    total_capacity = sum(capacity),
    median_capacity = median(capacity),
    cities = paste(sort(unique(city)), collapse = ", "),
    .groups = "drop"
  ) |>
  arrange(desc(total_capacity))

tract_capacity |>
  slice_head(n = 20)
```

```{r}
tract_capacity |>
  slice_max(total_capacity, n = 20) |>
  ggplot(aes(x = reorder(census_tract_geoid, total_capacity), y = total_capacity)) +
  geom_col(fill = "#31a354") +
  coord_flip() +
  labs(
    title = "Highest-capacity Census tracts",
    x = "Census tract GEOID",
    y = "Licensed capacity"
  ) +
  theme_minimal()
```

## Closed records

Use `date_closed` to keep closed records out of current-supply charts and to
summarize recently closed capacity for a separate audit. Treat closed capacity
as the facility's recorded licensed capacity, not as a confirmed seat-loss
measure for any particular program year.

```{r}
closed_capacity <- preschools_geo |>
  filter(is_closed, !is.na(date_closed)) |>
  mutate(closed_year = as.integer(format(date_closed, "%Y"))) |>
  group_by(city, closed_year) |>
  summarise(
    closed_facilities = n(),
    closed_capacity = sum(capacity, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(desc(closed_year), desc(closed_capacity))

closed_capacity
```

```{r}
ggplot(closed_capacity, aes(x = closed_year, y = closed_capacity, group = city)) +
  geom_line(color = "#636363") +
  geom_point(color = "#de2d26") +
  facet_wrap(vars(city)) +
  labs(
    title = "Closed capacity records by city and year",
    x = "Closure year",
    y = "Recorded capacity on closed facilities"
  ) +
  theme_minimal()
```

## Export report tables

Save both summary tables and row-level records. The summaries are useful for
reporting; the row-level table is what you need when a city, ZCTA, or tract
looks surprising.

```{r}
readr::write_csv(city_capacity, "capacity_by_city.csv")
readr::write_csv(zcta_capacity, "capacity_by_zcta.csv")
readr::write_csv(tract_capacity, "capacity_by_tract.csv")
readr::write_csv(preschools_geo, "capacity_analysis_rows.csv")
```
