---
title: "Capacity audits"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE,
  purl = FALSE
)
```

Use this workflow when you have a list of sites and need to compare your local
licensed-capacity records against the current CCLD facility record.

The examples are not evaluated during site builds because they call the live
CCLD Transparency API.

## Start with your site list

Your source file needs one column with CCLD license numbers and, if you want to
compare local values, one column with your local capacity value.

```{r}
library(ccldr)
library(dplyr)
library(readr)

sites <- read_csv("rr_sites_2026.csv", col_types = cols(.default = "c")) |>
  mutate(local_capacity = as.integer(local_capacity))
```

## Verify first

Run `ccld_verify()` first. It screens unknown or invalid licenses and returns
licensed capacity for every facility found by the API.

```{r}
verified <- ccld_verify(sites$license_number)

sites_verified <- sites |>
  bind_cols(verified |> select(-input))

sites_verified |>
  count(found, status, sort = TRUE)
```

`capacity` is parsed from the CCLD `CAPACITY` field. If a license is unknown,
invalid, or not found, `capacity` is `NA`.

## Compare capacities

Compare your local capacity value with the current CCLD value. Keep missing
values visible so they can be reviewed instead of silently dropping out.

```{r}
capacity_audit <- sites_verified |>
  mutate(
    capacity_delta = capacity - local_capacity,
    capacity_matches = !is.na(capacity) &
      !is.na(local_capacity) &
      capacity == local_capacity,
    needs_capacity_review = !found |
      is.na(capacity) |
      is.na(local_capacity) |
      capacity != local_capacity
  )
```

## Review mismatches

Sort records that need review by the size of the difference, then keep the
fields that make follow-up easy.

```{r}
capacity_audit |>
  filter(needs_capacity_review) |>
  arrange(desc(abs(capacity_delta)), site_name) |>
  select(
    site_name,
    license_number,
    found,
    facility_name,
    status,
    local_capacity,
    capacity,
    capacity_delta,
    date_closed
  )
```

If `capacity = NA` for a found facility, treat it as missing in the current
CCLD detail response and review the facility manually.

## Refresh deliberately

The package caches API responses while you work. For a capacity audit, clear
the cache at the start of a scheduled run or bypass the cache when checking one
important facility.

```{r}
ccld_cache_clear()

ccld_verify("013423996", cache = FALSE) |>
  select(facility_number, facility_name, status, capacity)
```

## A compact helper

For repeated audits, wrap the workflow in a small helper. This keeps the bulk
verification step explicit and makes the local capacity column configurable.

```{r}
add_capacity_audit <- function(data, license_col = license_number, capacity_col = local_capacity) {
  data <- data |>
    mutate(
      .license_input = as.character({{ license_col }}),
      .local_capacity = as.integer({{ capacity_col }})
    )

  verified <- ccld_verify(data$.license_input)

  data |>
    bind_cols(verified |> select(-input)) |>
    mutate(
      capacity_delta = capacity - .local_capacity,
      capacity_matches = !is.na(capacity) &
        !is.na(.local_capacity) &
        capacity == .local_capacity,
      needs_capacity_review = !found |
        is.na(capacity) |
        is.na(.local_capacity) |
        capacity != .local_capacity
    ) |>
    select(-.license_input)
}

capacity_audit <- add_capacity_audit(sites, license_number, local_capacity)
```
