---
title: "Closure date audits"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE,
  purl = FALSE
)
```

Use this workflow when you have a list of sites and need to know which ones
have closed, when they closed, and whether the CCLD record still matches your
local source.

The examples are not evaluated during site builds because they call the live
CCLD Transparency API.

## Start with your site list

Your source file only needs one column with CCLD license numbers. Keep whatever
site identifiers you use locally so the results can be joined back cleanly.

```{r}
library(ccldr)
library(dplyr)
library(readr)

sites <- read_csv("rr_sites_2026.csv", col_types = cols(.default = "c"))
```

Representative output:

```text
#> # A tibble: 3 x 2
#>   site_name                 license_number
#>   <chr>                     <chr>
#> 1 Johnson Family Child Care 13423996
#> 2 Ahmadi Family Child Care  013423958
#> 3 Unknown Site              99999999
```

## Verify first

Run `ccld_verify()` first. It screens unknown or invalid licenses and returns
current status and closure dates for every input row.

```{r}
verified <- ccld_verify(sites$license_number)

sites_verified <- sites |>
  bind_cols(verified |> select(-input))

sites_verified |>
  count(found, status, sort = TRUE)
```

Representative output:

```text
#> # A tibble: 3 x 3
#>   found status                         n
#>   <lgl> <chr>                      <int>
#> 1 TRUE  Licensed                     184
#> 2 TRUE  Closed, Licensee Initiated     6
#> 3 FALSE <NA>                           3
```

`ccld_verify()` is the right first pass for a full file. It includes
`capacity` from the CCLD `CAPACITY` field and `date_closed` from the
`DATECLOSED` field when the API has those values for the facility.

## Flag closed sites

Use `date_closed` and the current CCLD status to flag likely closures.

```{r}
closure_audit <- sites_verified |>
  mutate(
    is_closed = !is.na(date_closed) |
      grepl("^Closed", status, ignore.case = TRUE)
  )
```

Representative output:

```text
#> # A tibble: 3 x 6
#>   site_name                 facility_number found status                     date_closed is_closed
#>   <chr>                     <chr>           <lgl> <chr>                      <date>      <lgl>
#> 1 Johnson Family Child Care 013423996       TRUE  Licensed                   NA          FALSE
#> 2 Ahmadi Family Child Care  013423958       TRUE  Closed, Licensee Initiated 2026-04-17  TRUE
#> 3 Unknown Site              099999999       FALSE <NA>                       NA          FALSE
```

## Review closed and unmatched sites

Closed facilities usually have a `status` beginning with `"Closed"` and a
non-missing `date_closed`. Open facilities usually return `NA` for
`date_closed`.

```{r}
closure_audit |>
  filter(is_closed | !found) |>
  arrange(desc(is_closed), date_closed, site_name) |>
  select(
    site_name,
    license_number,
    found,
    facility_name,
    status,
    date_closed,
    last_visit_date
  )
```

Representative output:

```text
#> # A tibble: 2 x 7
#>   site_name                license_number found facility_name  status                     date_closed last_visit_date
#>   <chr>                    <chr>          <lgl> <chr>          <chr>                      <date>      <date>
#> 1 Ahmadi Family Child Care 013423958      TRUE  AHMADI, MARIAM Closed, Licensee Initiated 2026-04-17  2024-07-09
#> 2 Unknown Site             99999999       FALSE <NA>           <NA>                       NA          NA
```

If a facility is found but has `date_closed = NA`, treat it as not having a
closure date in the current CCLD detail response.

## Refresh deliberately

The package caches API responses while you work. For a closure audit, clear the
cache at the start of a scheduled run or bypass the cache when checking one
important facility.

```{r}
ccld_cache_clear()

closure_detail <- ccld_facility("013423958", cache = FALSE) |>
  select(facility_number, facility_name, status, date_closed)
```

Representative output:

```text
#> # A tibble: 1 x 4
#>   facility_number facility_name  status                     date_closed
#>   <chr>           <chr>          <chr>                      <date>
#> 1 013423958       AHMADI, MARIAM Closed, Licensee Initiated 2026-04-17
```

## A compact helper

For repeated audits, wrap the workflow in a small helper. This keeps the
bulk verification step explicit.

```{r}
add_closure_dates <- function(data, license_col = license_number) {
  data <- data |>
    mutate(.license_input = as.character({{ license_col }}))

  verified <- ccld_verify(data$.license_input)

  data |>
    bind_cols(verified |> select(-input)) |>
    mutate(
      is_closed = !is.na(date_closed) |
        grepl("^Closed", status, ignore.case = TRUE)
    ) |>
    select(-.license_input)
}

closure_audit <- add_closure_dates(sites, license_number)
```

Representative output:

```text
#> # A tibble: 3 x 5
#>   site_name                 facility_number found date_closed is_closed
#>   <chr>                     <chr>           <lgl> <date>      <lgl>
#> 1 Johnson Family Child Care 013423996       TRUE  NA          FALSE
#> 2 Ahmadi Family Child Care  013423958       TRUE  2026-04-17  TRUE
#> 3 Unknown Site              099999999       FALSE NA          FALSE
```
