---
title: "Closure date audits"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE,
  purl = FALSE
)
```

Use this workflow when you have a list of sites and need to know which ones
have closed, when they closed, and whether the CCLD record still matches your
local source.

The examples are not evaluated during site builds because they call the live
CCLD Transparency API.

## Start with your site list

Your source file only needs one column with CCLD license numbers. Keep whatever
site identifiers you use locally so the results can be joined back cleanly.

```{r}
library(ccldr)
library(dplyr)
library(readr)

sites <- read_csv("rr_sites_2026.csv", col_types = cols(.default = "c"))
```

## Verify first

Run `ccld_verify()` first. It screens unknown or invalid licenses and returns
current status and closure dates for every input row.

```{r}
verified <- ccld_verify(sites$license_number)

sites_verified <- sites |>
  bind_cols(verified |> select(-input))

sites_verified |>
  count(found, status, sort = TRUE)
```

`ccld_verify()` is the right first pass for a full file. It includes
`date_closed`, parsed from the CCLD `DATECLOSED` field, when the API has a
closure date for the facility.

## Flag closed sites

Use `date_closed` and the current CCLD status to flag likely closures.

```{r}
closure_audit <- sites_verified |>
  mutate(
    is_closed = !is.na(date_closed) |
      grepl("^Closed", status, ignore.case = TRUE)
  )
```

## Review closed and unmatched sites

Closed facilities usually have a `status` beginning with `"Closed"` and a
non-missing `date_closed`. Open facilities usually return `NA` for
`date_closed`.

```{r}
closure_audit |>
  filter(is_closed | !found) |>
  arrange(desc(is_closed), date_closed, site_name) |>
  select(
    site_name,
    license_number,
    found,
    facility_name,
    status,
    date_closed,
    last_visit_date
  )
```

If a facility is found but has `date_closed = NA`, treat it as not having a
closure date in the current CCLD detail response.

## Refresh deliberately

The package caches API responses while you work. For a closure audit, clear the
cache at the start of a scheduled run or bypass the cache when checking one
important facility.

```{r}
ccld_cache_clear()

closure_detail <- ccld_facility("013423958", cache = FALSE) |>
  select(facility_number, facility_name, status, date_closed)
```

## A compact helper

For repeated audits, wrap the workflow in a small helper. This keeps the
bulk verification step explicit.

```{r}
add_closure_dates <- function(data, license_col = license_number) {
  data <- data |>
    mutate(.license_input = as.character({{ license_col }}))

  verified <- ccld_verify(data$.license_input)

  data |>
    bind_cols(verified |> select(-input)) |>
    mutate(
      is_closed = !is.na(date_closed) |
        grepl("^Closed", status, ignore.case = TRUE)
    ) |>
    select(-.license_input)
}

closure_audit <- add_closure_dates(sites, license_number)
```
