---
title: "Troubleshooting and FAQ"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE,
  purl = FALSE
)
```

## Why did my license become 9 digits?

CCLD's Transparency API expects facility numbers in a 9-digit form. Many public
or local data files store the same number as 8 digits. `ccld_pad()` adds the
leading zero so requests use the API's canonical form.

```{r}
ccld_pad("13423996")
```

## Why does `ccld_verify()` return `found = FALSE`?

`ccld_verify()` keeps every input row. When the API does not return facility
detail for a license, the row stays in the result with `found = FALSE` and
missing facility fields. This makes failed matches visible after joins.

## Why did `ccld_pad()` error?

License inputs must contain only digits and must be 9 digits or fewer.

```{r}
ccld_pad("13423996")
ccld_pad(13423996)
```

Values such as `"ABC123"` or `"0123456789"` are rejected with a
`ccldr_invalid_input` error.

## Why are `"small_fccs"` and `"centers"` unsupported?

`ccld_alameda()` intentionally supports only API routes that return useful data:

- `"small_fccs"` is blocked by the API route used by the Transparency search.
- `"centers"` maps to a sparse legacy API bucket.

For small FCC enumeration, use the CKAN snapshot in
[`chekos/ccld-open-data-snapshot`](https://github.com/chekos/ccld-open-data-snapshot).
For center workflows, use `"preschools"`.

## What does the 250-result cap warning mean?

The CCLD search endpoint can cap individual responses at 250 records. To reduce
that risk, `ccld_alameda()` queries each Alameda city separately. If a city still
reports 250 or more results, the function warns because additional facilities
may be hidden by the API response cap.

## How do I force fresh data?

Use `cache = FALSE` for one call:

```{r}
ccld_verify("13423996", cache = FALSE)
```

Clear the whole cache before a reproducibility run:

```{r}
ccld_cache_clear()
```

Inspect cached responses:

```{r}
ccld_cache_info()
```

## How do I see API requests?

Set verbose mode before running a workflow:

```{r}
options(ccldr.verbose = TRUE)
```

Slow down batch requests when running larger jobs:

```{r}
options(ccldr.delay = 1)
```

## How should I run documentation checks locally?

Run the same core checks used for this package:

```r
roxygen2::roxygenise()
testthat::test_local(reporter = "summary")
pkgdown::build_site()
devtools::check(error_on = "never")
```

If a check fails only while executing live API examples, move that example into
a website-only article or make the chunk conditional.
